---
title: "Lab 10"
author: "Eugene Cheong"
date: "4/1/2020"
output: html_document
---

```{r, message= FALSE, echo=FALSE}
library(tidyverse)
library(maps)
library(mapdata)
library(lubridate)
library(viridis)
library(wesanderson)
library(ggplot2)
library(gganimate)
library(transformr)
library(plotly)
library(gifski)
library(DT)
```

## Methods 

This data was downloaded from te Johns Hopkins data site and sorted and knit together using R Studio 


```{r, echo=FALSE, message=FALSE}
daily_report <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/03-31-2020.csv")) %>% 
  rename(Long = "Long_")

```

## This is an example of the the data obtained from the Johns Hopkins repository
```{r, echo=FALSE, fig.height = 6, fig.width=6}
datatable(daily_report)
```


```{r, echo=FALSE, message=FALSE}
time_series_confirmed_long <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv")) %>%
  rename(Province_State = "Province/State", Country_Region = "Country/Region")  %>% 
               pivot_longer(-c(Province_State, Country_Region, Lat, Long),
                            names_to = "Date", values_to = "Confirmed") 

# Let's get the times series data for deaths

time_series_deaths_long <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv")) %>%
  rename(Province_State = "Province/State", Country_Region = "Country/Region")  %>% 
  pivot_longer(-c(Province_State, Country_Region, Lat, Long),
               names_to = "Date", values_to = "Deaths")

time_series_recovered_long <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_recovered_global.csv")) %>%
  rename(Province_State = "Province/State", Country_Region = "Country/Region") %>% 
  pivot_longer(-c(Province_State, Country_Region, Lat, Long),
               names_to = "Date", values_to = "Recovered")

# Create Keys 

time_series_confirmed_long <- time_series_confirmed_long %>% 
  unite(Key, Province_State, Country_Region, Date, sep = ".", remove = FALSE)

time_series_deaths_long <- time_series_deaths_long %>% 
  unite(Key, Province_State, Country_Region, Date, sep = ".") %>% 
  select(Key, Deaths)

time_series_recovered_long <- time_series_recovered_long %>% 
  unite(Key, Province_State, Country_Region, Date, sep = ".") %>% 
  select(Key, Recovered)

# Join tables
time_series_long_joined <- full_join(time_series_confirmed_long,
              time_series_deaths_long, by = c("Key"))

time_series_long_joined <- full_join(time_series_long_joined,
              time_series_recovered_long, by = c("Key")) %>% 
    select(-Key)

# Reformat the date
time_series_long_joined$Date <- mdy(time_series_long_joined$Date)

# Create Report table with counts
time_series_long_joined_counts <- time_series_long_joined %>% 
  pivot_longer(-c(Province_State, Country_Region, Lat, Long, Date),
               names_to = "Report_Type", values_to = "Counts")


```

## At a glance, we can see how quickly COVID-19 can spread and how some countries managed to flatten the curve
#

```{r, echo= FALSE, message= FALSE}
data_time <- time_series_long_joined %>% 
    group_by(Country_Region,Date) %>% 
    summarise_at(c("Confirmed", "Deaths", "Recovered"), sum) %>% 
    filter (Country_Region %in% c("China","Korea, South","Spain","US","Italy")) 

p <- ggplot(data_time, aes(x = Date,  y = Confirmed, color = Country_Region)) + 
      geom_point() +
      geom_line() +
      ggtitle("Confirmed COVID-19 Cases") +
      geom_point(aes(group = seq_along(Date))) +
      transition_reveal(Date) 
    
animate(p,renderer = gifski_renderer(), end_pause = 15)
```



## Number of Cases in the US. Hover mouse over each state to see how many cases there are in each
#

```{r, echo= FALSE, message= FALSE}
d_report <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/04-05-2020.csv")) %>% 
  rename(Long = "Long_") %>% 
  filter(Country_Region == "US") %>% 
  group_by(Province_State) %>% 
  summarize(Confirmed = sum(Confirmed)) %>% 
  mutate(Province_State = tolower(Province_State))

# load the US map data
us <- map_data("state")
# We need to join the us map data with our daily report to make one data frame/tibble
state_join <- left_join(us, d_report, by = c("region" = "Province_State"))

ggplotly(
  ggplot(data = us, mapping = aes(x = long, y = lat, group = group)) + 
    coord_fixed(1.3) + 
    # Add data layer
    geom_polygon(data = state_join, aes(fill = Confirmed), color = "black") +
    scale_fill_gradientn(colors = "light blue")+
    ggtitle("COVID-19 Cases in US") +
    theme(legend.position = "none")+
    # Cleaning up the graph
    labs(x=NULL, y=NULL) +
    theme(panel.border = element_blank()) +
    theme(panel.background = element_blank()) +
    theme(axis.ticks = element_blank()) +
    theme(axis.text = element_blank())
)

```


```{r, echo=FALSE, message=FALSE}

```

## Here is how Massachusetts looks like
```{r, echo=FALSE, message=FALSE}
daily_report <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/04-05-2020.csv")) %>% 
  rename(Long = "Long_") %>% 
  filter(Province_State == "Massachusetts") %>% 
  group_by(Admin2) %>% 
  summarize(Confirmed = sum(Confirmed)) %>% 
  mutate(Admin2 = tolower(Admin2))

us <- map_data("state")
ma_us <- subset(us, region == "massachusetts")
counties <- map_data("county")
ma_county <- subset(counties, region == "massachusetts")
state_join <- left_join(ma_county, daily_report, by = c("subregion" = "Admin2"))


# Making data entry for UMass 
group<- c(1184)
order<- c(37754)
region<- c("massachusetts")
subregion <- c("hampshire")
long <- c(-72.5301)
lat <- c(42.3868)

#Using rbind to add to original ma_county datatable
df <- rbind(ma_county, c(long, lat, group, order,region,subregion))
#Extracting UMass Datapoint in the most inconvenient way possible
b<- df[which.min(df$order), ]

#Making the annotation
a <- list(
  x = b$long,
  y = b$lat,
  text = "UMass Amherst",
  size = 4,
  xref = "x",
  yref = "y",
  showarrow = TRUE,
  arrowcolor = "maroon",
  arrowhead = 6,
  arrowsize = .5,
  ax = 30,
  ay = -75
)

fig1 <- ggplotly(
  ggplot(data = df, mapping = aes(x = long, y = lat, group = group)) + 
    coord_fixed(1.3) + 
    # Add data layer
    geom_polygon(data = state_join, aes(fill = Confirmed), color = "black") +
    scale_fill_gradientn(colours = 
                           wes_palette("Zissou1", 100, type = "continuous")) +
    ggtitle("COVID-19 Cases in MA") +
    # Cleaning up the graph
    labs(x= NULL, y= NULL) +
    theme(panel.border = element_blank()) +
    theme(panel.background = element_blank()) +
    theme(axis.ticks = element_blank()) +
    theme(axis.text = element_blank())
)

fig2 <- fig1 %>% layout(annotations = a)

fig2
```

## Confirmed Cases in Southeast Asian Countries
#### RIP my hometown of Malaysia. Hope the lockdown flattens the curve quickly
```{r, echo=FALSE, message=FALSE}
ts_SEA <- time_series_long_joined %>% 
  filter(Country_Region %in% c("Malaysia", "Singapore", "Indonesia", "Thailand", "Philippines", "Cambodia", "Vietnam", "Brunei", "Laos"))%>%
    ggplot(aes(x = Date, y = Confirmed, color= Country_Region)) + 
    geom_point() +
    geom_line() +
    facet_wrap(~Country_Region, ncol=3)+
  theme(axis.text.x = element_text(size = 7))+
  geom_point(aes(group = seq_along(Date))) +
  labs(title = 'Pandemic Day: {frame}', x = 'Date',y = 'Confirmed Cases')+
  transition_reveal(Date)
  
animate(ts_SEA, renderer = gifski_renderer(), end_pause = 20)

```


## References 

Johns Hopkins Data Repository [https://github.com/CSSEGISandData/COVID-19] 